---
title: "Open Policing Project Tutorial"
author: Chase Swenson
output:
  html_document:
    code_folding: hide
---


## Setup

First, let's load the necessary libraries and data that will allow us to
begin our investigation!

```{r setup, message=FALSE,  warning=FALSE}
## Libraries to include
library(tidyverse)
library(lubridate)

## Load the data
# Replace the path below with the path to where your data lives
data_path <- "https://datajournalism.tech/wp-content/uploads/2019/10/wichita.csv"
stops <- read_csv(data_path)

# Additional data and fixed values we'll be using
population_2016 <- tibble(
  subject_race = c(
    "asian/pacific islander", "black", "hispanic", "other/unknown","white"
  ),
  num_people = c(19294, 42485, 65090, 16686, 245499)
) %>% 
  mutate(subject_race = as.factor(subject_race))


center_lat <- 37.685260
center_lng <- -97.322562
```

## Exploratory Data Anyalysis
To have a quick look of the dataset, I willl run a summary of the data.

```{r}
colnames(stops)
nrow(stops)
str(stops)
summary(stops)
```

Here I will run a couple of pivot tables to inspect data in groups

```{r}
race_group <- stops %>% 
  group_by(subject_race) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
```




## Benchmark test

We saw police stopped white drivers more than black drivers. We need a baseline comparison to determine if the police are biased against one race.


### Stop rates



```{r}
population_2016 <- population_2016 %>% 
  mutate(prop = num_people / sum(num_people))
```

I will join the two data sets population_2016 and stops to see the actual ratio.

```{r}
stops_final <- stops %>% 
  count(subject_race) %>% 
  left_join(
    population_2016,
    by = "subject_race"
  ) %>% 
  mutate(stop_rate = n / num_people)
```

Second way fo computing actual stop rates

```{r}
stops_final <- race_group %>% 
  left_join(
    population_2016,
    by = "subject_race"
  ) %>% 
  mutate(stop_rate = n / num_people)
```

## Data VIsualization

```{r bar-chart}
#make the plot
bar <- ggplot(stops_final,
       aes(x=reorder(subject_race,stop_rate), y=stop_rate))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="red")+
  geom_hline(yintercept = 0) +
  labs(title="Stopped Drivers by Race",
       subtitle = "African American drivers got stopped more than White drivers in the city of Wichita, Kansas")+
  coord_flip()

options(scipen=10000)

bar
```

